---
title: "Data scraping"
author: "Nicholas Horton (nhorton@amherst.edu)"
date: "October 7, 2014"
output:
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---


```{r include=FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
```

```{r include=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).

# This changes the default colors in lattice plots.
trellis.par.set(theme=theme.mosaic())  

# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
```


```{r message=FALSE}
require(XML)
require(mosaic)
require(lubridate)
wikipedia = "http://en.wikipedia.org/wiki"
page = "List_of_nuclear_reactors"
result = readHTMLTable(paste(wikipedia, page, sep="/"),
  stringsAsFactors=FALSE)
table = result[[21]]  # change to appropriate table number
names(table)
```

```{r}
finaltable = mutate(table, 
      netcapacity = as.numeric(V6), 
      status = V5, 
      construct = dmy(V8))  # from lubridate
```


```{r}
tally(~ status, data=finaltable)
xyplot(netcapacity ~ construct, type=c("p", "smooth"), 
       main = "Capacity vs. construction date\nfor Japanese nuclear reactors", 
       data=finaltable)
```

Interpretation: the average net capacity of nuclear power plants in Japan tended to increase over time (but then plateaued in recent years).

### YOUR MISSION

In groups of 3, find an interesting Wikipedia page with a table, scrape the data from it, and generate a figure that tells an interesting story. (I'd suggest finding a simple table, as more complex ones are harder to parse.)

You should include a sentence which interprets the figure.

