
## Before

* Teach tab completion for variable names, function names, and arguments
* Using Rmd files.

### Regular Expressions

```{r eval=FALSE}
source("http://tinyurl.com/m4o4n2b/DCF/extract.R" )
```

Pull out 100,000 names from `BabyNames`, adding up the totals over the years by sex.  Call this `NameList`.  

* Names ending with "a".
* Names ending with a vowel.
* Names ending with a vowel or "y".
* Names with 3 consonants in a row.
* Names with 3 vowels in a row.

Interactive site for testing expressions: <http://regexone.com/>

#### Extraction

```{r eval=FALSE}
NameList %>% 
    extract(., "([aeiou])$", name, vowel=1 ) %>%
    group_by( sex, vowel ) %>% 
    summarise( total=sum(count) ) %>%
    arrange( sex, desc(total) )
```


## Starters

* [The janitorial problem](http://www.nytimes.com/2014/08/18/technology/for-big-data-scientists-hurdle-to-insights-is-janitor-work.html?_r=1) 
* Where will you get data in your own work.  Likely you'll have a connection.


### An example of available data

Files you can download: Examples:    
* US Census data <http://factfinder2.census.gov/faces/nav/jsf/pages/searchresults.xhtml?refresh=t>
* New York City 311 calls <https://data.cityofnewyork.us/Social-Services/311-Service-Requests-from-2010-to-Present/erm2-nwe9>.  Here's [the graph of complaint count against time](http://dtkaplan.github.io/DCF-2014/CourseNotes/ScrapingData/311-Calls.html) 

What variables are going to cause trouble?

## Reading directly from the web

[Nuclear plants activity](NuclearReactors.pdf)

Show the HTML markup for the [web page](http://en.wikipedia.org/wiki/List_of_nuclear_reactors) drawing attention to `<table>`


  They don't need to master this.  Conversion to CSV, yes. A tutorial? <http://www.r-bloggers.com/read-excel-files-from-r/>
  
### Big Data?

[Not so big activity](NotSoBig.pdf)



## Cleaning

* types: numerical, factor, string, ... how to examine them.
* are there values that just shouldn't be there: "J" for sex.
* checking against a priori bounds, e.g. month is 1:12, day is 1:31, wing-span is < 1 m
* inner_join against a translation table
* regular expressions using extract to, say, 
    * breaking a string with two or more variables into the separate variables.
    * pull out numbers from a string.  Using mutate() 
    * maybe [this package](http://www.r-bloggers.com/canned-regular-expressions-qdapregex-0-1-2-on-cran/) or <RegEx101.com>

UN City codes: <http://www.unece.org/fileadmin/DAM/cefact/locode/au.htm>

Activity: One about [birds](Notes/Week-3-Handout-2.pdf) 


