---
output:
  html_document:
    css: ~/KaplanFiles/DCF-Course-2014/CSS/DCF-Course-style.css
    fig_caption: yes
  pdf_document: default
---

```{r child="../../CSS/DCF-common.Rmd"}
```

## Demo: Wide and Narrow

```{r}
require(rworldmap)
data(countrySynonyms)
```

`View()` the `countrySynonyms` table.
Is this in wide or narrow format?  What would the other format look like?

To convert it to narrow format ... Hint: In the end, you may not need the "key" variable.

```{r}
require(tidyr)
Translate <- 
  countrySynonyms %>%
  select(-ID) %>%
  mutate( ISO3=toupper(ISO3)) %>%
  gather( value=country, key=bogus, name1:name8) %>%
  select( -bogus ) %>%
  mutate( country=ifelse(country %in% c("","<NA>"), NA, country)) %>%
  na.omit
```

Ideas about which is to be preferred?

## All together:

Population by ISO2 and ISO3

Country Codes at ...

<http://www.fao.org/countryprofiles/iso3list/en/>

Country Synonyms:
```{r}
Synonyms <- fetchData("DCF/CountrySynonyms.csv")
```

ISO2 and ISO3 codes
```{r}
Codes <- fetchData("DCF/CountryCodes.csv")
```


1. Pull out the country and population data from `CountryData`
#. Translate `country` in `CountryData` to `ISO3` using `Synonyms`
#. Then translate `ISO3` to `ISO2` using `Codes`
#. Merge with `WorldCities` by `country`/`ISO2`


## World Cities

Check the table for plausibility.  
* What's the total number of people represented?
```{r}
WorldCities %>% 
  summarise( total=sum(population))
```

* How many cities larger than 100,000? Than 1,000,000?

* Make a scatterplot of the latitude and longitude of the cities larger than 100K
```{r}
p <-  
  WorldCities %>%
  filter( population > 1e5 ) %>%
  ggplot( aes(x=longitude, y=latitude )) +
    geom_point(alpha=.1, aes(size=population)) +
    scale_size_area()
```

* What's the biggest city in each country?
```{r}
BiggestByCountry <- 
  WorldCities %>% 
  group_by( country ) %>%
  filter( rank(desc(population)) == 1 )
```

Add that to the plot as another layer.  Make it red.
```{r}
p + geom_point( data=BiggestByCountry, color='red')
```

* Countries where the biggest city is more than 5M people
```{r}
BiggestByCountry %>%
  filter( population > 5000000)
```

Connecting with population ...

## In Groups

Working with your group, figure out a strategy for adding in, for each country, the total population of that country.




## Give them `CountryByISO2`







* Country name synonyms


* Cities that are the largest, by fraction of the population, in each country.
    * Need the country population data: from `CountryData`
    ```{r}
Population <-
  CountryData %>%
  select( country, pop )
```

Now join these to the `WorldCities` data ...

`WorldCities` has ISO-2 country codes.

`CountryData` had common names (in English).  We need to translate from one to the other.

A table is available via
```{r}
Codes <- fetchData("DCF/CountryCodes.csv")
```

Pull out the country and ISO-2 variables
```{r}
Codes <-
  Codes %>%
  select(country, ISO2)
```

Join with `Population`
```{r}
Population <-
  Population %>%
  inner_join( Codes )
```

Add  `ISO2` column to `WorldCities` in place of `country`
```{r}
WorldCities <-
  WorldCities %>%
  mutate( ISO2=country) %>%
  select( -country )
```

Merge `Population` with `WorldCities` on `ISO2`
```{r eval=FALSE}
WorldCities <-
  WorldCities %>%
  inner_join( Population )
```



World cities:

The basic questions ...

* Northernmost, Southermost Closest to the equator
* Fraction of country population that lives in cities.
* Countries where the biggest city is twice as large as the next biggest.
* Compare the pop of the biggest city to the other cities.  Find the rank of the city that is less than half the population of the biggest cities.




## Medicare

Generate city-wise population, age, race statistics.

For a given DRG, or body-system, look at per-capita number of procedures as function of age.   

How many of the US cities are in the World Cities data table? 

Find the cities that don't have a match.

For each city in the US, find the area of that city, the population density etc.

Plot out the old-age cities.

Plot out the cities 

For each DRG, find the three top and bottom states in terms of number of procedures per capita.

Graphing networks: just a demo

Long and narrow: just a demo

# Global migration -- Everyone together

http://data.worldbank.org/data-catalog/global-bilateral-migration-database

Countries that send people to the US.  Make a map.

Trends over time of countries that send people to the US.  (Convert to narrow, use `geom_lines()` with `group` property.)

Set color by GDP of the sending country.

Bilateral migration --- put the inbound and outbound migration as different variables.  A network based on ...

* Balance between countries
* What's total migration as a fraction of the population?
* Which countries have strong gender imbalance of migration?

### Rank

Within each state, pull out the 100 zip codes with the highest minority populations.  Plot these out on a choropleth map of overall minority population.  (Or, maybe, foreign born.)

# Class Notes: Week 4

Drill on wide/narrow:  
* pivot, fold, gather, spread, unite, ...
* NCI60 --- just the probe data.  Change to narrow format and find the probes with the highest standard deviation of expression across all cell lines.

Baby names: put the boy count side by side with the girl count.  Then find the names that are given to both boys and girls.

### Rank

Within each state, pull out the 100 zip codes with the highest minority populations.  Plot these out on a choropleth map of overall minority population.  (Or, maybe, foreign born.)

You're doing the background research for a birdwatching manual.  

* In the bird data, for each month, pull out the 4 highest species.
* Pull out the species that are in the top 20 every month.  These are the year-round birds.
* Modify the 4 highest in each-month to exclude the year-round birds.  (This will be an anti-join.

For each country, pull out the two countries that have the strongest bilateral immigration.  Make a network graph of these

### Long and Narrow

Turn the immigration data into narrow, so that you can calculate the pairs that are close to being bilateral.  

Doing a transform on a grouped variable allows you, if you want, to use aggregate quantities for that group.  For instance, the groups of males and females have different mean heights.  Here's how you would calculate the difference of an individual from his or her group's mean.

```{r eval=FALSE}
NHANES <- group_by( NHANES, sex )
NHANES <- mutate( NHANES, dheight=height-mean(height) )
summarise( NHANES, sd(dheight) )
```

## Still to do

Do a tabulation of the rank-choice data.

Find the variances of the NCI data, by probe.

Cross-tab of NHANES: perhaps cardiac death versus sex or BMI. 




Using `ggplot()` directly.  Layers ...

## Constructing the Chart

Consider three different grammar components of the charts:

1. Frame
2. Glyphs, known as *geoms* in `ggplot2`.
3. Facets

There must always be a frame and at least one glyph layer in a (non-empty) chart.

Write down an appropriate `ggplot2` statement for generating each component. Say what aesthetics apply to each layer.  (Any layer can inherit aesthetics from the frame, or can override them.)  Take `TheData` to be a data frame containing relevant variables.  If other dataframes are needed, call them `Data1`, `Data2`, etc.

1. Frame: `ggplot( data = TheData,  ? )`
#. Layer 1: 
#. Layer 2:
#. Layer 3:
#. Faceting:

Write down an appropriate command to construct the graph (or something like it).  As a reminder, here are a few of the `ggplot2` functions:

`ggplot()`, `geom_point()`, `geom_barchart()`, `geom_blank()`, `geom_boxplot()`, `geom_density()`, `geom_errorbar()`, `geom_polygon()`, `geom_line()`, `geom_path()`, `geom_ribbon()`, `geom_rug()`, `geom_text()`, `scale_y_log10()`, `scale_x_log10()`, `xlab()`, `ylab()`, `ggtitle()`

