---
title: "DCF Week 7 Warm-Up"
output:
  pdf_document:
    fig_height: 3
    fig_width: 5
    keep_tex: no
  html_document:
    css: ~/KaplanFiles/DCF-Course-2014/CSS/DCF-Course-style.css
    fig_caption: no
  word_document:
    fig_height: 3
    fig_width: 5
date: "Data and Computing Fundamentals"
---
```{r}
require(rpart)
```

Tonight is about review and adding an important new approach to interpreting data.

We have spent most of our time on two subjects:

1. making data-oriented graphics
2. transfiguring data: getting from the data we are given to the "glyph-ready" data that you need to make a graphic or some other mode to guide our interpretation of the data.

* Problems from Stanford/CMC intro computer science course.
* A new data set to explore with graphics
* If time, a clustering approach to genetics.

Then we'll go on to machine learning.  Think of it as a way of presenting data that allows us to handle multiple variables.

What is learning?  Recognizing patterns knowing facts.  Gutenberg Bible, 1492, Reformation

Perhaps "understanding" is broader: knowing where to look for patterns and how to apply a kind of pattern to a new situation.

Machines can learn: they can spot patterns.

Supervised and Unsupervised.



1. Two problems from Stan



## Unsupervised Learning

### Cluster the countries

### NCI 60

### Show a glaucoma prediction

```{r eval=FALSE}
data("GlaucomaM", package = "TH.data")
glaucoma_rpart <- rpart(Class ~ ., data = GlaucomaM, 
                        control = rpart.control(xval = 100)
```

### Calculate a diabetes prediction

What are the highest risk groups?

## Supervised Learning

### Baby Names

Look at all the letters in girls' names, compare to boy's names.

As features, use length, last letter, 

### Predictors of Immigration






* Build a decision-tree model of the factors that determine strong one-way emigration.  (Ratio of GDP, life expectancy, ...)

