---
output:
  pdf_document:
    keep_tex: yes
    toc: yes
  html_document:
    css: ../../CSS/DCF-style.css
    fig_caption: yes
    toc: yes
---

```{r child="../../CSS/DCF-common.Rmd"}
```

```{r include=FALSE}
require(xtable)
```



## Birds

Use the `OrdwayBirds` and `OrdwaySpeciesNames` data tables to   
* find the peak month for each major species
* find major species that are seen in good numbers for at least 6 months of the year
* identify, for the major species over each month, whether they tend to be seen almost only at the beginning or end of the month.

```{r echo=FALSE,message=FALSE}
Corrected <-
  OrdwayBirds %>%
  inner_join( OrdwaySpeciesNames ) %>% 
  select( Species=SpeciesNameCleaned, Month, Day ) %>%
  na.omit() ## cleaned 
MaxMonth <- Corrected %>%
  group_by( Species, Month ) %>%
  summarise( n=n() ) %>% 
  filter( rank( desc(n)) < 2 )
MajorSpeciesIDs <-
  Corrected %>%
  group_by( Species ) %>% 
  summarise( n=n() ) %>%
  filter( rank( desc(n)) < 10 )
OnlyMajorSpecies <-
  Corrected %>% 
  inner_join( MajorSpeciesIDs %>% select(Species)) %>%
  droplevels( ) # drop the non-major species levels
BigMonths <-
  OnlyMajorSpecies %>%
  group_by( Month, Species ) %>%
  summarise( n=n() ) %>%
  filter( n > 0.02*sum(n) )
SixMonthsOrMore <-
  BigMonths %>% 
  group_by( Species ) %>%
  summarise( months=n() ) %>% 
  filter( months >= 6 )
TimeOfMonth <-
  OnlyMajorSpecies %>% 
  inner_join( select(MaxMonth,-n)) %>%
  mutate( Day=as.numeric(Day) ) %>%
  group_by( Month, Species ) %>%
  summarise( early=mean(Day<=10),
             mid = mean(Day>10 & Day<20),
             late= mean(Day>=20))
```



## Disasters





Case studies: Disasters.  Which forms cause the most damage, greatest loss of life, economic loss.

* Rates in the country data, e.g. miles of road per area, 
* Fraction alive, cardio, other in NHANES

### Filter

WEEK 4 -- CVC/Summer2014/Activities/Migration.Rmd --- there are all sorts of rank comparisons here involving the maximum, top five, and so on.



Pick out the zip codes you would target for:

* lots of elderly.
* young people and high income.

* Where do the elderly live in the US?  Zip-code data: see <file:///Users/kaplan/KaplanFiles/MOSAIC/CVC/Summer2014/Notes/GraphicsBasics/graphics-basics.html#17>



Medicare spending: see the questions at <file:///Users/kaplan/KaplanFiles/DCF-2014/CaseStudies/MedicareSpending/notes.html>

* Groupwise means and sd.
* Counts using group_by() and n().  (Group by the levels of the variable to be counted, perhaps including another variable for subdivisions.)
* Converting counts to proportions. (Divide by sum().)
    ```{r}
    mosaicData::KidsFeet %>% group_by( biggerfoot,sex ) %>% 
    summarise(count=n() )
    ```
* Graphics with mPlot(), mBar(), mMap()

* Add a layer to a map, perhaps zip-code population with state boundaries.

* Who votes for whom?  Wade county voters: create cross tabs on registration, etc.  Make bar charts to display patterns.


## Week 3

EXERCISE:

Find the location of 

* the most northern and most southern city.  The city closest to 
* the equator.

END OF EXERCISE.
