---
output:
  html_document:
    css: ~/KaplanFiles/DCF-Course-2014/CSS/DCF-Course-style.css
    fig_caption: yes
  pdf_document: default
---

```{r child="../../CSS/DCF-common.Rmd"}
```


# As it played out that night ...

### Agenda

* Review the [command guide](../../CourseNotes/Guides/EnoughDCF.pdf) 
    You should try to reach a point where the contents make sense.  You don't have to master every command; use the sheet for a reference.
  
  
* Basic strategy: know your input and imagine your output.  Think in English about 
    * what data verbs to use
    * the arguments to each verb
* Some `BabyNames` exercises, talking through it and having each student implement the commands:
    * How many kids are there? (about 333M)  `summarise()`
    * What are the five most popular names over all the years. `group_by()`, `summarise()`, `arrange()` `desc()` `head()`
    * What are the five most popular names in 2013? `filter()` `arrange()` `desc()`
    * What are the two most popular names for each year?  What are the two most popular in each year for each sex? `group_by()` `summarise()` `rank()` `<= 2`
    
* Making a small Rmd file
    * Open from template.
    * Save with an appropriate name.
    * Change the "This file.Rmd" in the boilerplate. Don't screw with the YAML.
    * Compile.  Did it work?  No, check the previous step.
    * Delete the instruction text from the template.
    * Add some text: "These are the most popular baby names..."
    * Add a chunk: The 5 most popular baby names.
    * Compile to HTML
    * Upload the HTML to moodle.


* The [data verb worksheet](../../Background/DataVerb-Worksheet.pdf) How the flow chart works.
    * Make the graph of names over the years --- they can choose their own names.
    
    
# Class Notes: Week 3

## All together now

### Command Guide
Review the 1-page command guide: ["Enough R for DCF"](http://dtkaplan.github.io/DCF-2014/CourseNotes/Guides/EnoughDCF.pdf)



### World Cities data.

* Plot the latitude and longitude of the cities. --- already in glyph-ready form.
* Find the total urban population of each country.  
    * Plot this as a choropleth map.
    * Plot this as a scatter plot, using median latitude and longitude for each country to set position.
* Extremes
    * Largest city in each country.
    * Find the northernmost city. 
    * Find the northernmost city in each country with a population > 100000
    * Find the city closest to the equator.

## Names

* Find the top 5 in each year.  Note that the grouping still holds after a filter.
```{r}
Top5 <-
  BabyNames %>% 
  group_by( year, sex ) %>%
  filter( rank(desc(count))<=5 )
groups( Top5 )
```

* Find those names that have remained in the top five for more than 20 years.
```{r}
LongRunning <- 
  Top5 %>% ungroup() %>%
  group_by( name, sex ) %>%
  filter( n() > 20 ) %>% 
  summarize( nyears=n() )
```

* Of these, find the 10 female and 10 male names with the shortest run (closest to 20 years)
```{r}
OurGroup <- 
  LongRunning %>%
  group_by( sex ) %>%
  filter( rank(nyears)<=10)
```

## Small Group Activities

1. [Popularity of names](Notes/Week-3/Week-3-Handout.pdf)
2. [Bird species](Notes/Week-3/Week-3-Handout.pdf)


WEEK 4 -- CVC/Summer2014/Activities/Migration.Rmd --- there are all sorts of rank comparisons here involving the maximum, top five, and so on.



Pick out the zip codes you would target for:

* lots of elderly.
* young people and high income.

* Where do the elderly live in the US?  Zip-code data: see <file:///Users/kaplan/KaplanFiles/MOSAIC/CVC/Summer2014/Notes/GraphicsBasics/graphics-basics.html#17>



Medicare spending: see the questions at <file:///Users/kaplan/KaplanFiles/DCF-2014/CaseStudies/MedicareSpending/notes.html>

* Groupwise means and sd.
* Counts using group_by() and n().  (Group by the levels of the variable to be counted, perhaps including another variable for subdivisions.)

* Add a layer to a map, perhaps zip-code population with state boundaries.

* Who votes for whom?  Wade county voters: create cross tabs on registration, etc.  Make bar charts to display patterns.



EXERCISE:



END OF EXERCISE.
